---
title: "p8105_hw1_sm5592"
author: "Shaolei Ma"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, echo = F, message = F, results = 'hide'}
# load necessary packages
library(moderndive)
library(ggplot2)
library(tidyverse)
Sys.setlocale("LC_TIME", "US") # change default time zone
```

# Problem 1
First, load the data set and describe it.
```{r}
data("early_january_weather") # load the dataset
summary(early_january_weather$temp)
```

 * The data set has `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.  
 * It contains the following 15 variables: `r variable.names(early_january_weather)`. As the temperature and humidity are about to be evaluated below, I take these as important variables. For temperature, the `min`, `max`, `median` values respectively are `r min(early_january_weather$temp)`, `r max(early_january_weather$temp)`, `r median(early_january_weather$temp)`; For humidity,  the `min`, `max`, `median` values respectively are `r min(early_january_weather$humid)`, `r max(early_january_weather$humid)`, `r median(early_january_weather$humid)`. 
 
Then, make a scatterplot.
```{r, message = F}
ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) + 
  geom_point() + labs(title = "Scatterplot of Temperature ~ Time") +
  theme(text = element_text(family = "serif")) # change fonts
ggsave("scatterplot_hw1.jpg")
```

From the scatterplot, the temperature increases overall over the time period and fluctuates within each day with the peak forming usually in the afternoon. The humidity is relatively steady except for the time window between January 13 and January 14, and the humidity drop on January 14 marks the highest temperature over the period.

# Problem 2
Create a data frame comprised of:  

 * a random sample of size 10 from a standard Normal distribution  
 * a logical vector indicating whether elements of the sample are greater than 0  
 * a character vector consisting of the sample converted to character  
 * a factor vector measuring the value of the sample

```{r}
dat = 
  tibble(
    var_numeric = rnorm(10),
    var_logical = var_numeric > 0,
    var_char = as.character(var_numeric),
    var_factor = factor(case_when(
      var_numeric < 0 ~ "1", # if x < 0, then label it "low".
      var_numeric >= 0 & var_numeric < 0.5 ~ "2", # if 0 <= x < 0.5, then label it "med".
      var_numeric >= 0.5 ~ "3"), # if x >= 0.5, then label it "high".
      labels = c("low", "med", "high"),
      levels = c("1", "2", "3"), ordered = T
      )
  )
```

Show the first rows of the data frame: 
```{r, echo = F}
head(dat)
```

Calculate the means of each variable:
```{r}
for(variable in variable.names(dat))
  print(paste("The mean for variable '", variable, "' is ", dat %>% pull(variable) %>% mean %>% round(digits = 2), ".", sep = ""))
```

Both the numeric and logical variables yield results, while the character and factor variables yield NA.  
Try to convert the logical, character, and factor variables to numeric variables.
```{r, results = 'hide'}
for(variable in variable.names(dat))
{
  if(variable == "var_numeric") next # skip the numeric variable
  print(paste("The mean for variable '", variable, "' is ", dat %>% pull(variable) %>% as.numeric %>% mean %>% round(digits = 2), ".", sep = ""))
}
```

All yield an output. It could be concluded that for the function mean(), the logical variable is converted to numeric and then computed, while the character and factor variables are not converted by default.